Staged Abstract Interpreters

### 2019-09-09 paper/80_related.tex
typos
polish

\paragraph{Partial Evaluation and Multi-stage Programming}
Partial evaluation is an automatic technique for program
specialization and was studied comprehensively by
\citet{10.1007/3-540-61580-6_11, DBLP:books/daglib/0072559}.
In this paper, we use mutli-stage programming to specialize programs.
The Lightweight Modular Staging (LMS) framework
\cite{DBLP:conf/gpce/RompfO10} we used in the paper relies on
type-level stage annotations.  Other notable MSP systems using syntactic
quotations exist in the ML family, e.g., MetaML \cite{DBLP:conf/pepm/TahaS97}
and MetaOCaml \cite{DBLP:conf/gpce/CalcagnoTHL03, DBLP:conf/flops/Kiselyov14}. 
Compared with other approaches using syntactic quotations, LMS provides
guarantee for preserving evaluation orders within a stage.
\citet{Ofenbeck:2017:SGP:3136040.3136060} and
\citet{Amin:2017:CTI:3177123.3158140} discussed staging for generic programming
as well as stage polymorphism in LMS.
%The idea of staging an abstract interpreter we presented here
%is also applicable to other MSP systems.
Multi-stage programming has been widely used to improve the
performance in many other domains, such as optimizing compilers or
domain-specific languages \cite{DBLP:conf/pldi/RompfSBLCO14,
  DBLP:conf/snapl/RompfBLSJAOSKDK15,
  DBLP:journals/tecs/SujeethBLRCOO14, DBLP:conf/gpce/SujeethGBLROO13,
  DBLP:journals/jfp/CaretteKS09}, numerical computation \cite{PGL-038,
  DBLP:conf/pepm/AktemurKKS13}, generic programming
\cite{DBLP:journals/pacmpl/Yallop17,
  Ofenbeck:2017:SGP:3136040.3136060}, data processing
\cite{DBLP:conf/oopsla/JonnalageddaCSRO14,
  DBLP:conf/popl/KiselyovBPS17}, query compilation in databases
\cite{DBLP:conf/osdi/EssertelTDBOR18, DBLP:conf/sigmod/TahboubER18},
etc.

As a source of inspiration of this paper, the Futamura projections reveal a
close relation between interpreters and compilers, which was
originally proposed by \citet{futamura1971partial}, and later reprinted \cite{Futamura1999}.
\citeauthor{Amin:2017:CTI:3177123.3158140} considered a tower of concrete
interpreters and showed how to collapse them by using MSP -- it would be interesting
to explore this idea for multiple abstract interpreters
\cite{Cousot:2019:AAI:3302515.3290355, Giacobazzi:2015:APA:2676726.2676987}.
\citet{10.1007/11561347_18} discussed a monad with let-insertion and
memoization for code generation, using Gaussian Elimination in MetaOCaml as an
example. In this work, let-insertion is handled by the ANF transformation in
LMS, and the memoizing monad is just a combination of a reader monad and a state monad
operating on staged data.  Similar to the idea in this paper, specializing
monadic interpreters was explored by \citet{danvy1991compiling,
DBLP:conf/dsl/SheardBP99}. However, they are different from the performance motivation
here: the generated code is in monadic form.

>>>>

\paragraph{Partial Evaluation and Multi-stage Programming}
Partial evaluation is an automatic technique for program
specialization and was studied comprehensively by
\citet{10.1007/3-540-61580-6_11, DBLP:books/daglib/0072559}.
In this paper, we use mutli-stage programming to specialize programs.
The Lightweight Modular Staging (LMS) framework
\cite{DBLP:conf/gpce/RompfO10} we used in the paper relies on
type-level stage annotations.  Other notable MSP systems using syntactic
quotations exist in the ML family, e.g., MetaML \cite{DBLP:conf/pepm/TahaS97}
and MetaOCaml \cite{DBLP:conf/gpce/CalcagnoTHL03, DBLP:conf/flops/Kiselyov14}. 
Compared with other approaches using syntactic quotations, LMS provides
guarantees for preserving evaluation orders within a stage.
\citet{Ofenbeck:2017:SGP:3136040.3136060} and
\citet{Amin:2017:CTI:3177123.3158140} discussed staging for generic programming
as well as stage polymorphism in LMS.
%The idea of staging an abstract interpreter we presented here
%is also applicable to other MSP systems.
Multi-stage programming has been widely used to improve the
performance in many other domains, such as optimizing compilers or
domain-specific languages \cite{DBLP:conf/pldi/RompfSBLCO14,
  DBLP:conf/snapl/RompfBLSJAOSKDK15,
  DBLP:journals/tecs/SujeethBLRCOO14, DBLP:conf/gpce/SujeethGBLROO13,
  DBLP:journals/jfp/CaretteKS09}, numerical computation \cite{PGL-038,
  DBLP:conf/pepm/AktemurKKS13}, generic programming
\cite{DBLP:journals/pacmpl/Yallop17,
  Ofenbeck:2017:SGP:3136040.3136060}, data processing
\cite{DBLP:conf/oopsla/JonnalageddaCSRO14,
  DBLP:conf/popl/KiselyovBPS17}, query compilation in databases
\cite{DBLP:conf/osdi/EssertelTDBOR18, DBLP:conf/sigmod/TahboubER18},
etc.

As a source of inspiration of this paper, the Futamura projections reveal a
close relation between interpreters and compilers, which was
originally proposed by \citet{futamura1971partial}, and later reprinted \cite{Futamura1999}.
\citeauthor{Amin:2017:CTI:3177123.3158140} considered a tower of concrete
interpreters and showed how to collapse them by using MSP -- it would be interesting
to explore this idea for multiple abstract interpreters
\cite{Cousot:2019:AAI:3302515.3290355, Giacobazzi:2015:APA:2676726.2676987}.
\citet{10.1007/11561347_18} discussed a monad with let-insertion and
memoization for code generation, using Gaussian Elimination in MetaOCaml as an
example. In this work, let-insertion is handled by the ANF transformation in
LMS, and the memoizing monad is just a combination of a reader monad and a state monad
operating on staged data.  Similar to the idea in this paper, specializing
monadic interpreters was explored by \citet{danvy1991compiling,
DBLP:conf/dsl/SheardBP99}. However, they are different from the performance
motivation here.
